The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-19 paper/mri_paper.tex
nondimensionalization, thin gap

\newcommand\reye{\mathrel{Re}}
\newcommand\reym{\mathrel{Rm}}

@@

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space. In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.

@@

\begin{figure}[h!]
\centering
\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow. We investigate a 2D slice of the X-Z (radial-vertical) plane. Our domain is the bolded black box, of width L. The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
\end{figure}

@@

\beq
thin gap equation,
\eeq

the width of the channel is small compared to its distance from the center of rotation. The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}). Implications of this approximation are discussed further in \S\ref{discussion}.

@@

We perturb the fluid quantities with three-dimensional, axisymmetric perturbations. We nondimensionalize the equations according to [...]
The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the perturbed quantities.

>>>>

\newcommand\reye{\mathrm{Re}}
\newcommand\reym{\mathrm{Rm}}

@@

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space. In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.

@@

\begin{figure}[h!]
\centering
\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow. We investigate a 2D slice of the X-Z (radial-vertical) plane. Our domain is represented by the bolded black box, of width L. The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
\end{figure}

@@

\beq
\left(R_2 - R_1\right) \ll \frac{1}{2}\left(R_2 + R_1\right),
\eeq

where $R_1$ is the radius of the inner cylinder and $R_2$ is the radius of the outer cylinder, such that the width of the channel is small compared to its distance from the center of rotation. The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}). Implications of this approximation are discussed further in \S\ref{discussion}.

@@

We perturb the fluid quantities with three-dimensional, axisymmetric perturbations. We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{setup}). We define the nondimensional Reynolds number,

\beq
\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu},
\eeq

and magnetic Reynolds number,

\beq
\reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}.
\eeq

The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the nondimensional, perturbed quantities.
